---
title: "A published chapter that runs on your computer"
subtitle: "Static Quarto HTML · active R chunks · your local CarmaR engine"
---

This is a proof of concept for publishing a normal `.qmd` chapter with active
R chunks. The page itself is static and hosted on GitHub Pages. When you press
**Run**, the visible code executes in the R installation on your computer and
the fresh result returns to this page.

## Run locally

Install the current CarmaR app once. Then press **Run on my computer** or any
chunk's **Run** button. The browser may ask once whether this site may connect
to software on your local network; choose **Allow**.

CarmaR starts without a helper tab and the status above the chunks changes to
**CarmaR connected**. Nothing executes until you press **Run**.

## 1. Create a dataset

This first chunk creates data in your local R session. The code is editable;
change it directly before running if you wish.

```{r}
#| eval: false
set.seed(2026)
study <- data.frame(
  learner = paste0("L", 1:12),
  group = rep(c("practice", "comparison"), each = 6),
  score_before = round(rnorm(12, 62, 7), 1)
)
study$score_after <- round(
  study$score_before + ifelse(study$group == "practice", 9, 3) + rnorm(12, 0, 2),
  1
)
study
```

## 2. Continue in the same R session

Cells share state. This chunk uses `study`, which was created above.

```{r}
#| eval: false
aggregate(
  cbind(score_before, score_after) ~ group,
  data = study,
  FUN = mean
)
```

## 3. Produce a real local plot

The PNG is drawn by your installed R—not by JavaScript or webR—and returned to
the published chapter.

```{r}
#| eval: false
cols <- ifelse(study$group == "practice", "#2859c5", "#e28a2b")
plot(
  study$score_before,
  study$score_after,
  pch = 19,
  col = cols,
  xlab = "Score before",
  ylab = "Score after",
  main = "Learning gains"
)
abline(0, 1, lty = 2, col = "grey50")
legend(
  "topleft",
  legend = c("Practice", "Comparison"),
  col = c("#2859c5", "#e28a2b"),
  pch = 19,
  bty = "n"
)
```

## 4. Fit a model with base R

No optional packages are needed for this proof.

```{r}
#| eval: false
fit <- lm(score_after ~ score_before + group, data = study)
coef(summary(fit))
```

## What this proves

- The source remains an ordinary Quarto `.qmd` file.
- The published result remains static HTML suitable for GitHub Pages.
- Every visible R chunk becomes a real CarmaR editor with **Run** and **Stop**.
- Cells execute sequentially in one local R session.
- CarmaR's console, data-frame, plot, and export components render inline.
- The native app authorizes only this exact publishing origin for this session.
- No extension, helper tab, AI panel, or per-chunk permission is involved.

R code is not sandboxed: inspect unfamiliar chunks before running them.
