---
title: "A published chapter that runs on your computer"
subtitle: "Static Quarto HTML · active R chunks · your local CarmaR engine"
---

This is a proof of concept for publishing a normal `.qmd` chapter with active
R chunks. The page itself is static and hosted on GitHub Pages. When you press
**Run**, the visible code executes in the R installation on your computer and
the fresh result returns to this page.

## Start local CarmaR

Run these two lines once in your R console:

```r
source("https://mohsaqr.github.io/carmar-runnable-book-poc/install.R")
```

The installer starts or reuses the correct local engine. If an older CarmaR is
already running, it replaces that stale process so the pairing route cannot
return `not found`.

Then press **Run on my computer** above the first active chunk. CarmaR opens a
small local consent window naming this site. Keep it open while using the
chapter. Nothing executes until you press a chunk's **Run** button.

## 1. Create a dataset

This first chunk creates data in your local R session. Click **Edit** if you
want to change it before running.

```{r}
#| eval: false
set.seed(2026)
study <- data.frame(
  learner = paste0("L", 1:12),
  group = rep(c("practice", "comparison"), each = 6),
  score_before = round(rnorm(12, 62, 7), 1)
)
study$score_after <- round(
  study$score_before + ifelse(study$group == "practice", 9, 3) + rnorm(12, 0, 2),
  1
)
study
```

## 2. Continue in the same R session

Cells share state. This chunk uses `study`, which was created above.

```{r}
#| eval: false
aggregate(
  cbind(score_before, score_after) ~ group,
  data = study,
  FUN = mean
)
```

## 3. Produce a real local plot

The PNG is drawn by your installed R—not by JavaScript or webR—and returned to
the published chapter.

```{r}
#| eval: false
cols <- ifelse(study$group == "practice", "#2859c5", "#e28a2b")
plot(
  study$score_before,
  study$score_after,
  pch = 19,
  col = cols,
  xlab = "Score before",
  ylab = "Score after",
  main = "Learning gains"
)
abline(0, 1, lty = 2, col = "grey50")
legend(
  "topleft",
  legend = c("Practice", "Comparison"),
  col = c("#2859c5", "#e28a2b"),
  pch = 19,
  bty = "n"
)
```

## 4. Fit a model with base R

No optional packages are needed for this proof.

```{r}
#| eval: false
fit <- lm(score_after ~ score_before + group, data = study)
coef(summary(fit))
```

## What this proves

- The source remains an ordinary Quarto `.qmd` file.
- The published result remains static HTML suitable for GitHub Pages.
- Every visible R chunk receives **Edit** and **Run** controls.
- Cells execute sequentially in one local R session.
- Console output, data frames, errors, warnings, and plots return inline.
- The public site never receives direct access to the loopback WebSocket.

R code is not sandboxed: inspect unfamiliar chunks before running them. The
local bridge disconnects when you close its small CarmaR window or stop the
kernel.
